NumPy---Array Programming in Python

### 2020-01-26 body.tex
Work on outline and ecosystem section

To manipulate the data structure in an intuitive manner, we provide users with
a simple yet powerful array expression syntax.
It supports operations such as indexing (selecting subarrays) and elementwise
calculations.
A high level syntax allows users to express themselves succinctly, while NumPy
deals with the underlying mechanics of making operations fast.

@@

\section*{Supporting the ecosystem}

\fixme{Maybe some wording around the actual *support* element?}

@@

Users predominantly interact with NumPy arrays using {\em indexing} (to access
subarrays or individual elements), {\em operators} (e.g., $+$, $-$, $\times$,
and, for matrix multiplication, $@$), as well as {\em array-aware functions};
together, these provide an easily readable, expressive, high-level API for
array programming.

Various libraries build on and extend this functionality to further enrich the
ecosystem of available tools.  For example, SciPy adds various mathematical,
scientific, and engineering routines; Pandas provides indexed arrays, known as
dataframes; xarray implements arrays with labeled axes; and Matplotlib
generates publication-ready figures and visualizations.  Employing these
low-level array-aware functions are widely used technique-specific libraries,
including scikit-learn (machine learning), scikit-image (image processing),
statsmodels (statistics),  networkx (graphs and complex networks)
\cite{SciPyProceedings_11}, and Napari (volumetric image visualization).

@@

Exposing array programming primitives, as well as the surrounding ecosystem of
tools, in Python---an interpreted language---creates an ideal environment for
interactive, exploratory data analysis where users may iteratively inspect,
manipulate, and visualize their data.

IPython \cite{perez2007ipython}

Jupyter \cite{Kluyver:2016aa}

\section*{The proliferation of arrays}

\fixme{Add references somewhere to Zarr, https://arrow.apache.org/,
https://parquet.apache.org/,
https://github.com/ruby-numo/numo-narray/wiki/Comparison-with-Numpy}

>>>>

Users predominantly interact with NumPy arrays using {\em indexing} (to access
subarrays or individual elements), {\em operators} (e.g., $+$, $-$, $\times$,
and, for matrix multiplication, $@$), as well as {\em array-aware functions};
together, these provide an easily readable, expressive, high-level API for
array programming, while NumPy
deals with the underlying mechanics of making operations fast.

@@

The expressive power and brevity of NumPy's array programming syntax combined with 
an advanced interactive environment like IPython \cite{perez2007ipython}
or Jupyter \cite{Kluyver:2016aa} make interactive data exploration fluid and simple.

\section*{Scientific Python ecosystem}

@@

While NumPy provides a solid foundation for array programming, the scientific
Python ecosystem is composed of many different libraries, which build on and
extend this functionality.
These libraries can roughly be organized into three layers:
core libraries, technique-specific libraries, and domain-specific libraries.

The core libraries include SciPy, Matplotlib, Pandas, and xarray.
SciPy adds various mathematical, scientific, and engineering routines;
Matplotlib generates publication-ready figures and visualizations,
Pandas provides indexed arrays, known as dataframes; and
xarray implements arrays with labeled axes.

Employing these low-level array-aware functions are widely used
technique-specific libraries, including scikit-learn (machine learning),
scikit-image (image processing), statsmodels (statistics),  networkx (graphs
and complex networks) \cite{SciPyProceedings_11}, and Napari (volumetric image
visualization).

@@

% we should have an example of how these tools may all be used in an
% example analysis.  maybe one of the ones mentioned in the abstract?

Exposing array programming primitives, as well as the surrounding ecosystem of
tools, in Python---an interpreted language---creates an ideal environment for
interactive, exploratory data analysis where users may iteratively inspect,
manipulate, and visualize their data.
The rich ecosystem of tools building on NumPy and the larger community
of scientists and programmers using those tools mean that NumPy has a
responsibility to maintain considerable stability.
This need creates interesting challenges for NumPy development as the
computational environment scientists are analyzing their data on continues
to evolve and change.

\section*{Proliferation of arrays}

%\fixme{Add references somewhere to Zarr, https://arrow.apache.org/,
%https://parquet.apache.org/,
%https://github.com/ruby-numo/numo-narray/wiki/Comparison-with-Numpy}
